Capabilities, MMIO and Robust Safety

### 2020-04-16 capability-notes/driver-notes/main.tex
Few small tweaks + auctex master file

It should be noted that a more general, semantic version of this specification
is also possible. It would contain the following elements, that should all
follow directly from the specification above (TODO: write this out properly and
do the same for the semantic bootspec):
\begin{itemize}
\item All registers except \X{PC} are in the value relation with respect to
  either the empty world $\Wemp$ (if we do not want to allow any non-empty, non-enter
  capabilities in registers), or a world $\Winit$ containing a single standard region for
  all of the adversary's memory.
  The \X{PC} register satisfies both the read- and write conditions (note the
  parallel with the fundamental theorem) with respect to this same world, if we
  picked the non-empty world, or just points to the adversary's memory, which
  does not contain capabilities (if we
  picked the empty world, a capability pointing into memory can never be valid).
  It feels slightly less artificial to pick the non-empty world option here.
\item We have possession of the predicates $\I{region}$ and
  $\I{sts\_full\_world}$ for the world we picked in the previous bullet. we will
  need these to apply the fundamental theorem.
\item A description of the exact driver code, still in terms of points-to chunks
  (or, in this case equivalently, an invariant). We can write this invariant as
      $\driverC(b,e,\overline{\V{w}_l}) = \knowInv{\iota}{\bigast{l \in [b,e)}{l \mapsto_{a} \mathit{\V{w}_l}}}$, with $\V{w}_l$
      some hard-coded driver word (an instruction or a capability) at location $l$.
\end{itemize}

@@

Notice how the syntactic check for the registers is only performed on the set of
registers $r$, allowing the closures in $\overline{r_{a_i}}$ to point into to
the driver address space (TODO: we should change this though and add a
requirement to $\initOKo$).

@@

  \X{bootSpec}(P, g, b, e, A_{\X{entry}},\V{bootCont}) \eqdef \\[0.8em]
    \quad
  \begin{array}{l}
    \left\{
    \begin{array}{l}
      \X{PC} \mapsto_{r} (\X{RWX}, \X{G}, 0, \X{MEM_{MAX}}, l_{\X{boot}}) \ast {} \\[0.5em]
      \bigast{r \in \X{Reg} \setminus \X{PC}}{r \mapsto_{r} r_{0}[r]}
      \ast \bigast{m \in \X{Mem} \setminus \MMIO}{m \mapsto_{a} m_{0}[m]}
      \ast \; \tracefrag{\emptyset} \ast \\[1.2em]
      \later  \V{bootCont}
    \end{array}
    \right\}
    \\[2.5em]
    \quad \I{Seq}\; (\I{Instr}\; \I{Executable})
    \\[0.2em]
    \left\{ \Phi \right\}
  \end{array}\\[4em]

@@

 \X{bootCont_{sem}}(P, g, b, e, A_{\X{entry}},\Phi,\Winit) \eqdef  \\[0.2em]
    \begin{array}{l}
       (\exists b_{\X{adv}}\; e_{\X{adv}}\; a_{\X{adv}}\; w_r.\; \\[0.5em]
      \quad
      \begin{array}{l}
        \X{PC} \mapsto_{r} (\X{RWX}, \X{G}, b_{\X{adv}}, e_{\X{adv}}, a_{\X{adv}})
        \ast \VR^{\notMMIO} (\Winit)(\X{RW}, \X{G}, b_{\X{adv}}, e_{\X{adv}}, a_{\X{adv}})) \ast {} \\[1.2em]
        \bigast{r \in Reg\setminus \{\X{PC}\}}{(r \mapsto_{r} w_r \ast
        \VR^{\notMMIO} (\Winit)(w_r))} \ast {} \\[1.2em]
        \driverC(b,e,\overline{\V{w}_l})\ast
        \; \trInv \; \ast \\[0.5em]
        \X{region}(\Winit) \ast  \X{sts\_full\_world}\; \X{sts\_std} \;
        \Winit)\\[0.5em]       \end{array}
    \end{array} \\[4em]
    \quad\wand
    \begin{array}{l}
     \wpre{\I{Seq}\; (\I{Instr}\; \I{Executable})}{\Phi}
    \end{array}\\[1.2em]

  \X{bootSpec_{sem}}(P, g, b, e, A_{\X{entry}},\Winit)
    \eqdef\\
     \quad \forall \Phi\ldotp  \X{bootSpec}(P, g, b, e, A_{\X{entry}},\X{bootCont_{sem}}(P, g, b, e, A_{\X{entry}},\Phi,\Winit))\\[0.8em]

@@

\begin{mathpar}
  \mprset{vskip=0.5em}
  \inferrule
  {  \forall a_{i} \in A_{\X{entry}} \ldotp\,\left\{
    {\begin{array}{ll}
    \trInv \ast
    \driverC(b,e,\overline{\V{w}_l})\!
         \vdash \\
         \forall W \Wpriv \Winit \ldotp\;\ER^{\notMMIO} (W) (\X{RX}, g, b, e, a_i)
        \end{array}} \right. \\   \X{bootSpec}^{\X{sem}}(P,g,b,e,A_{\X{entry}}, \Winit) \\
  (r_0, m_0, \emptyset) \longrightarrow^* (r, m, t)\\
  }
  {P(t)}
\end{mathpar}

@@

One could also take as a parameter a predicate that restrict the set of traces
that are allowed (that is what~\cite{sammler-2020} do).

@@

>>>>

It should be noted that a more general, semantic version of this specification
is also possible. It would contain the following elements, that should all
follow directly from the specification above:
\begin{itemize}
\item All registers except \X{PC} are in the value relation with respect to
  either the empty world $\Wemp$ (if we do not want to allow any non-empty, non-enter
  capabilities in registers), or a world $\Winit$ containing a single standard region for
  all of the adversary's memory.
  The \X{PC} register satisfies both the read- and write conditions (note the
  parallel with the fundamental theorem) with respect to this same world, if we
  picked the non-empty world, or just points to the adversary's memory, which
  does not contain capabilities (if we
  picked the empty world, a capability pointing into memory can never be valid).
  It feels slightly less artificial to pick the non-empty world option here.
\item We have possession of the predicates $\I{region}$ and
  $\I{sts\_full\_world}$ for the world we picked in the previous bullet. we will
  need these to apply the fundamental theorem.
\item A description of the exact driver code, still in terms of points-to chunks
  (or, in this case equivalently, an invariant). We can write this invariant as
      $\driverC(b,e,\overline{\V{w}_l}) = \knowInv{\iota}{\bigast{l \in [b,e)}{l \mapsto_{a} \mathit{\V{w}_l}}}$, with $\V{w}_l$
      some hard-coded driver word (an instruction or a capability) at location $l$.
\end{itemize}

@@

Notice how the syntactic check for the registers is only performed on the set of
registers $r$, allowing the closures in $\overline{r_{a_i}}$ to point into to
the driver address space (TODO: we should change this though and add a
requirement to $\initOKo$). (TODO 2: initOK\_sem is simply the precondition of bootcont\_sem below, we should inline it here or delete it)

@@

  \X{bootSpec}(P, g, b, e,\V{bootCont}) \eqdef \\[0.8em]
    \quad
  \begin{array}{l}
    \left\{
    \begin{array}{l}
      \X{PC} \mapsto_{r} (\X{RWX}, \X{G}, 0, \X{MEM_{MAX}}, l_{\X{boot}}) \ast {} \\[0.5em]
      \bigast{r \in \X{Reg} \setminus \X{PC}}{r \mapsto_{r} r_{0}[r]}
      \ast \bigast{m \in \X{Mem} \setminus \MMIO}{m \mapsto_{a} m_{0}[m]}
      \ast \; \tracefrag{\emptyset} \ast \\[1.2em]
      \later  \V{bootCont}
    \end{array}
    \right\}
    \\[2.5em]
    \quad \I{Seq}\; (\I{Instr}\; \I{Executable})
    \\[0.2em]
    \left\{ \Phi \right\}
  \end{array}\\[4em]

@@

 \X{bootCont_{sem}}(P, g, b, e, \Phi,\Winit) \eqdef  \\[0.2em]
    \begin{array}{l}
       (\exists b_{\X{adv}}\; e_{\X{adv}}\; a_{\X{adv}}\;\overline{w_r}.\; \\[0.5em]
      \quad
      \begin{array}{l}
        \X{PC} \mapsto_{r} (\X{RWX}, \X{G}, b_{\X{adv}}, e_{\X{adv}}, a_{\X{adv}})
        \ast \VR^{\notMMIO} (\Winit)(\X{RW}, \X{G}, b_{\X{adv}}, e_{\X{adv}}, a_{\X{adv}})) \ast {} \\[1.2em]
        \bigast{r \in Reg\setminus \{\X{PC}\}}{(r \mapsto_{r} w_r \ast
        \VR^{\notMMIO} (\Winit)(w_r))} \ast {} \\[1.2em]
        \driverC(b,e,\overline{\V{w}_l})\ast
        \; \trInv \; \ast \\[0.5em]
        \X{region}(\Winit) \ast  \X{sts\_full\_world}\; \X{sts\_std} \;
        \Winit)\\[0.5em]       \end{array}
    \end{array} \\[4em]
    \quad\wand
    \begin{array}{l}
     \wpre{\I{Seq}\; (\I{Instr}\; \I{Executable})}{\Phi}
    \end{array}\\[1.2em]

  \X{bootSpec_{sem}}(P, g, b, e, \Winit)
    \eqdef\\
     \quad \forall \Phi\ldotp  \X{bootSpec}(P, g, b, e,\X{bootCont_{sem}}(P, g, b, e, \Phi,\Winit))\\[0.8em]

@@

\begin{mathpar}
  \mprset{vskip=0.5em}
  \inferrule
  {  \forall a_{i} \in A_{\X{entry}} \ldotp\,\left\{
    {\begin{array}{ll}
    \trInv \ast
    \driverC(b,e,\overline{\V{w}_l})\!
         \vdash \\
         \forall W \Wpriv \Winit \ldotp\;\ER^{\notMMIO} (W) (\X{RX}, g, b, e, a_i)
        \end{array}} \right. \\   \X{bootSpec}^{\X{sem}}(P,g,b,e, \Winit) \\
  (r_0, m_0, \emptyset) \longrightarrow^* (r, m, t)\\
  }
  {P(t)}
\end{mathpar}

@@

One could also take as a parameter a predicate that restricts the set of traces
that are allowed (that is what~\cite{sammler-2020} do).

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
